Through the Eyes of the Teacher - Multimodal Exploration of Expertise Differences in the Perception of Classroom Disruptions in a Laboratory Study

### 2024-11-26 02_paper/02_study/02_expertise_paper/expertise_2024_11_18_MK.Rmd
x paper mwu test

##################### DURATION OF DISRUPTIONS ########################

# 1. Data Preparation: Filter and Calculate Disruption Duration
df_aoi_disrup <- df_aoi %>%
  filter(TOI %in% disruptions, Duration_of_interval <= 30000) %>%
  select(Group, Participant, TOI, Duration_of_interval) %>%
  mutate(
    Group = as_factor(Group),
    Disruption_Category = case_when(
      TOI %in% verbal_disruptions ~ "Verbal disruptions",
      TOI %in% physical_disruptions ~ "Physical disruptions",
      TOI %in% lack_of_eagerness ~ "Lack of eagerness to learn",
      TRUE ~ NA_character_
    ),
    Duration_of_interval_sec = round(Duration_of_interval / 1000, 2)
  ) %>%
  filter(!is.na(Disruption_Category)) %>% # Keep only rows with valid categories
  drop_na()

@@

# 2. Descriptive Statistics by Group (Overall)
summary_stats <- df_aoi_disrup %>%
  group_by(Group) %>%
  summarise(
    Mean = round(mean(Duration_of_interval_sec, na.rm = TRUE), 2),
    SD = round(sd(Duration_of_interval_sec, na.rm = TRUE), 2),
    .groups = "drop"
  )

# 3. Descriptive Statistics by Group and Category
disruption_stats_table <- df_aoi_disrup %>%
  group_by(Group, Disruption_Category) %>%
  summarise(
    N = n(),
    Mean = round(mean(Duration_of_interval_sec, na.rm = TRUE), 2),
    SD = round(sd(Duration_of_interval_sec, na.rm = TRUE), 2),
    Min = round(min(Duration_of_interval_sec, na.rm = TRUE), 2),
    Max = round(max(Duration_of_interval_sec, na.rm = TRUE), 2),
    .groups = "drop"
  )

# Overall t-Test and Cohen's d for Plot Annotations
t_test_overall <- t.test(Duration_of_interval_sec ~ Group, data = df_aoi_disrup, var.equal = TRUE)
d_overall <- cohen.d(Duration_of_interval_sec ~ Group, data = df_aoi_disrup)

# By Category t-Test and Cohen's d for Plot Annotations
category_stats <- df_aoi_disrup %>%
  group_by(Disruption_Category) %>%
  summarise(
    t_value = t.test(Duration_of_interval_sec ~ Group, var.equal = TRUE)$statistic,
    df = t.test(Duration_of_interval_sec ~ Group, var.equal = TRUE)$parameter,
    p_value = round(t.test(Duration_of_interval_sec ~ Group, var.equal = TRUE)$p.value, 3),
    cohen_d = round(cohen.d(Duration_of_interval_sec ~ Group)$estimate, 2),
    .groups = "drop"
  )

@@

# Plot 1: Disruption Duration by Group (Overall)
y_max <- max(df_aoi_disrup$Duration_of_interval_sec) * 1.5

plot_duration_group <- ggplot(df_aoi_disrup, aes(x = Group, y = Duration_of_interval_sec, fill = Group)) +
  geom_boxplot(outlier.shape = NA, width = 0.5) +
  geom_jitter(width = 0.15, alpha = 0.5, color = "black", size = 1.5) +
  labs(x = "Group", y = "Disruption Duration (seconds)") +
  theme_cowplot() +
  scale_fill_brewer(palette = "RdBu") +
  ggtitle("Disruption Duration by Group") +
  theme(legend.position = "none") +
  ylim(0, y_max) +
  annotate("text", x = 1, y = y_max * 0.85,
           label = paste0("M = ", summary_stats$Mean[summary_stats$Group == "Novice"],
                          "\nSD = ", summary_stats$SD[summary_stats$Group == "Novice"]),
           hjust = 0.5, size = 4, fontface = "italic") +
  annotate("text", x = 2, y = y_max * 0.85,
           label = paste0("M = ", summary_stats$Mean[summary_stats$Group == "Expert"],
                          "\nSD = ", summary_stats$SD[summary_stats$Group == "Expert"]),
           hjust = 0.5, size = 4, fontface = "italic") +
  annotate("text", x = 1.5, y = y_max * 0.95,
           label = paste0("p = ", format.pval(t_test_overall$p.value, digits = 3, eps = .001),
                          "\nd = ", round(d_overall$estimate, 2)),
           hjust = 0.5, size = 4, fontface = "italic")

@@

# Plot 2: Disruption Duration by Group and Category
plot_duration_category <- ggplot(df_aoi_disrup, aes(x = Group, y = Duration_of_interval_sec, fill = Group)) +
  geom_boxplot(outlier.shape = NA, width = 0.5) +
  geom_jitter(width = 0.15, alpha = 0.5, color = "black", size = 1.5) +
  facet_wrap(~Disruption_Category, scales = "free", labeller = label_wrap_gen(width = 15)) +
  labs(x = "Group", y = "Disruption Duration (seconds)") +
  theme_cowplot() +
  scale_fill_brewer(palette = "RdBu") +
  ggtitle("Disruption Duration by Group and Category") +
  theme(legend.position = "none") +
  ylim(0, y_max) +
  geom_text(data = category_stats, aes(
    x = 1.5, y = y_max * 0.85,
    label = paste0("p = ", p_value, "\nd = ", cohen_d)
  ), inherit.aes = FALSE, size = 4, fontface = "italic")

@@

##################### t-TEST AND EFFECT SIZE RESULTS ########################

# 1. Overall t-Test and Cohen's d
t_test_overall <- t.test(Duration_of_interval_sec ~ Group, data = df_aoi_disrup, var.equal = TRUE)
d_overall <- cohen.d(Duration_of_interval_sec ~ Group, data = df_aoi_disrup)

# 2. t-Test and Cohen's d Results by Category
category_stats <- df_aoi_disrup %>%
  group_by(Disruption_Category) %>%
  summarise(
    t_value = round(t.test(Duration_of_interval_sec ~ Group, var.equal = TRUE)$statistic, 2),
    df = round(t.test(Duration_of_interval_sec ~ Group, var.equal = TRUE)$parameter, 0),
    p_value = round(t.test(Duration_of_interval_sec ~ Group, var.equal = TRUE)$p.value, 3),
    cohen_d = round(cohen.d(Duration_of_interval_sec ~ Group)$estimate, 2),
    "95% CI (d)" = paste0(
      "[", round(cohen.d(Duration_of_interval_sec ~ Group)$conf.int[1], 2),
      ", ", round(cohen.d(Duration_of_interval_sec ~ Group)$conf.int[2], 2), "]"
    ),
    .groups = "drop"
  )

# 3. Format Overall t-Test Results for APA Table
t_test_result_overall <- data.frame(
  "t-value" = round(t_test_overall$statistic, 2),
  "df" = round(t_test_overall$parameter, 0),
  "p-value" = round(t_test_overall$p.value, 3),
  "Mean Difference" = round(diff(t_test_overall$estimate), 2),
  "Cohen's d" = round(d_overall$estimate, 2),
  "95% CI (d)" = paste0("[", round(d_overall$conf.int[1], 2), ", ", round(d_overall$conf.int[2], 2), "]")
)

# Display Overall t-Test Results in APA Style
kable(
  t_test_result_overall,
  caption = "t-Test and Effect Size for Disruption Duration (Overall)",
  row.names = FALSE
) %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)

# 4. Display Category-Specific t-Test Results in APA Style
kable(
  category_stats,
  caption = "t-Test and Effect Size for Disruption Duration by Category",
  col.names = c("Category", "t-value", "df", "p-value", "Cohen's d", "95% CI (d)"),
  row.names = FALSE
) %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)

>>>>

# Prepare data for analysis: filter, calculate, and categorize disruptions
df_aoi_disrup <- df_aoi %>%
  filter(TOI %in% disruptions, Duration_of_interval <= 30000) %>%
  select(Group, Participant, TOI, Duration_of_interval) %>%
  mutate(
    Group = as_factor(Group),
    Disruption_Category = case_when(
      TOI %in% verbal_disruptions ~ "Verbal disruptions",
      TOI %in% physical_disruptions ~ "Physical disruptions",
      TOI %in% lack_of_eagerness ~ "Lack of eagerness to learn",
      TRUE ~ NA_character_
    ),
    Duration_of_interval_sec = round(Duration_of_interval / 1000, 2)
  ) %>%
  filter(!is.na(Disruption_Category)) %>%
  drop_na()

@@

# Descriptive Statistics by Group
summary_stats_group <- df_aoi_disrup %>%
  group_by(Group) %>%
  summarise(
    N = n(),
    Mean = round(mean(Duration_of_interval_sec, na.rm = TRUE), 2),
    SD = round(sd(Duration_of_interval_sec, na.rm = TRUE), 2),
    Median = round(median(Duration_of_interval_sec, na.rm = TRUE), 2),
    IQR = round(IQR(Duration_of_interval_sec, na.rm = TRUE), 2),
    Min = round(min(Duration_of_interval_sec, na.rm = TRUE), 2),
    Max = round(max(Duration_of_interval_sec, na.rm = TRUE), 2),
    .groups = "drop"
  )

# Descriptive Statistics by Category and Group
summary_stats_category_group <- df_aoi_disrup %>%
  group_by(Disruption_Category, Group) %>%
  summarise(
    N = n(),
    Mean = round(mean(Duration_of_interval_sec, na.rm = TRUE), 2),
    SD = round(sd(Duration_of_interval_sec, na.rm = TRUE), 2),
    Median = round(median(Duration_of_interval_sec, na.rm = TRUE), 2),
    IQR = round(IQR(Duration_of_interval_sec, na.rm = TRUE), 2),
    Min = round(min(Duration_of_interval_sec, na.rm = TRUE), 2),
    Max = round(max(Duration_of_interval_sec, na.rm = TRUE), 2),
    .groups = "drop"
  )

# Display Descriptive Tables
kable(
  summary_stats_group,
  caption = "Descriptive Statistics for Disruption Duration by Group",
  col.names = c("Group", "N", "Mean", "SD", "Median", "IQR", "Min", "Max"),
  row.names = FALSE
) %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)

kable(
  summary_stats_category_group,
  caption = "Descriptive Statistics for Disruption Duration by Category and Group",
  col.names = c("Category", "Group", "N", "Mean", "SD", "Median", "IQR", "Min", "Max"),
  row.names = FALSE
) %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)

@@

# Plot 1: Disruption Duration by Group
plot_duration_group <- ggplot(df_aoi_disrup, aes(x = Group, y = Duration_of_interval_sec, fill = Group)) +
  geom_boxplot(outlier.shape = NA, width = 0.5) +
  geom_jitter(width = 0.15, alpha = 0.5, color = "black", size = 1.5) +
  labs(x = "Group", y = "Disruption Duration (seconds)") +
  theme_cowplot() +
  scale_fill_brewer(palette = "RdBu") +
  ggtitle("Disruption Duration by Group") +
  theme(legend.position = "none") +
  annotate("text", x = 1, y = max(df_aoi_disrup$Duration_of_interval_sec) + 4,
           label = paste0("Median = ", summary_stats_group$Median[summary_stats_group$Group == "Novice"],
                          "\nIQR = ", summary_stats_group$IQR[summary_stats_group$Group == "Novice"]),
           hjust = 0.5, size = 4) +
  annotate("text", x = 2, y = max(df_aoi_disrup$Duration_of_interval_sec) + 4,
           label = paste0("Median = ", summary_stats_group$Median[summary_stats_group$Group == "Expert"],
                          "\nIQR = ", summary_stats_group$IQR[summary_stats_group$Group == "Expert"]),
           hjust = 0.5, size = 4)

@@

# Plot 2: Disruption Duration by Group and Category
plot_duration_category <- ggplot(df_aoi_disrup, aes(x = Group, y = Duration_of_interval_sec, fill = Group)) +
  geom_boxplot(outlier.shape = NA, width = 0.5) +
  geom_jitter(width = 0.15, alpha = 0.5, color = "black", size = 1.5) +
  facet_wrap(~Disruption_Category, scales = "free", labeller = label_wrap_gen(width = 15)) +
  labs(x = "Group", y = "Disruption Duration (seconds)") +
  theme_cowplot() +
  ylim(0, 30) + 
  scale_fill_brewer(palette = "RdBu") +
  ggtitle("Disruption Duration by Group and Category") +
  theme(legend.position = "none")

@@

# Normality Tests (Shapiro-Wilk)
normality_tests <- df_aoi_disrup %>%
  group_by(Group) %>%
  summarise(
    Shapiro_Wilk_p = round(shapiro.test(Duration_of_interval_sec)$p.value, 3),
    .groups = "drop"
  )

kable(
  normality_tests,
  caption = "Shapiro-Wilk Test for Normality by Group",
  col.names = c("Group", "Shapiro-Wilk p-value"),
  row.names = FALSE
) %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)

# Mann-Whitney U-Test and Effect Size
mann_whitney_test <- wilcox.test(Duration_of_interval_sec ~ Group, data = df_aoi_disrup, exact = FALSE)

# Effect Size Calculation for Mann-Whitney (r)
n1 <- nrow(df_aoi_disrup %>% filter(Group == "Novice"))
n2 <- nrow(df_aoi_disrup %>% filter(Group == "Expert"))
r_effect <- abs(mann_whitney_test$statistic - (n1 * n2 / 2)) / sqrt((n1 * n2 * (n1 + n2 + 1)) / 12)

mann_whitney_results <- data.frame(
  "Test" = "Mann-Whitney U",
  "W" = round(mann_whitney_test$statistic, 2),
  "p-value" = round(mann_whitney_test$p.value, 3),
  "Effect Size (r)" = round(r_effect, 3)
)

kable(
  mann_whitney_results,
  caption = "Mann-Whitney U-Test Results for Disruption Duration",
  row.names = FALSE
) %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)

# Kruskal-Wallis Test and Pairwise Comparisons
kruskal_test <- kruskal.test(Duration_of_interval_sec ~ Disruption_Category, data = df_aoi_disrup)

kruskal_results <- data.frame(
  "Test" = "Kruskal-Wallis",
  "Chi-squared" = round(kruskal_test$statistic, 2),
  "df" = kruskal_test$parameter,
  "p-value" = round(kruskal_test$p.value, 3)
)

kable(
  kruskal_results,
  caption = "Kruskal-Wallis Test Results for Differences Between Disruption Categories",
  row.names = FALSE
) %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)

# Pairwise Wilcoxon test with Bonferroni correction
pairwise_wilcox <- pairwise.wilcox.test(
  x = df_aoi_disrup$Duration_of_interval_sec,
  g = df_aoi_disrup$Disruption_Category,
  p.adjust.method = "bonferroni"
)

# Extract and round the p-values to 3 decimal places
pairwise_results <- as.data.frame(pairwise_wilcox$p.value) %>%
  mutate(across(everything(), ~ round(., 3)))

# Display the results with rounded p-values
kable(
  pairwise_results,
  caption = "Pairwise Wilcoxon Test Results (Bonferroni Corrected)",
  row.names = TRUE
) %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)
